---
title: "data cleanning"
author: "Chen Zhou"
date: "9/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(readxl)
library(dplyr)
library(ggplot2)
df <- read_excel("Practicum 1.xlsx")

```

```{r}
table(is.na(df))
```

```{r}
# compute change in BMI(kg/m^2)

# compute starting weight
df$start_weight_lbs = df$bweight - df$pounds_gained

# convert lbs to kg
df$start_weight_kg = 0.453592 * df$start_weight_lbs

# convert inch to meter
df$height_meter = 0.0254 * df$height

# compute starting BMI 
df$BMI_start = df$start_weight_kg/(df$height_meter^2)

# compute change in BMI
df$BMI_change = df$BMI - df$BMI_start
```

```{r}
# imputation for Total_met_min
df$Total_met_min_imp = 8* df$`Vig ex Time` + 4* df$`Mod ex time` + 3.3*df$`Walk ex Time`
```

```{r}
sum(is.na(df$Age))
sum(is.na(df$Total_met_min_imp))
sum(is.na(df$BMI_change))
```

```{r}
library(arsenal)
newdf <- df[c("gender", "Age", "shift", "Total_met_min_imp", "BMI_change")]
summary(newdf, title="old")
data_complete_cases <- newdf[complete.cases(newdf), ]
summary(data_complete_cases)
head(data_complete_cases)
```

```{r}
sum(is.na(data_complete_cases$Age))
sum(is.na(data_complete_cases$Total_met_min_imp))
sum(is.na(data_complete_cases$BMI_change))
```


### Exploratory data analysis

```{r}
ggplot(data_complete_cases,aes(y = log(BMI_change), x = log(Total_met_min_imp), color = gender) )+
geom_point() + geom_smooth(method = "lm", fill = NA) + 
  xlab("log(Total Metabolic Minutes)") + 
  ylab("log(Change in BMI)") + ggtitle("The relationship between BMI change and Total metabolic minutes by gender")
```
Based on the scatterplot of the relationship between Total Metabolic Minutes and BMI change by gender, we log both the Total Metabolic Minutes and Change in BMI. The result shows that female has a smaller slope than male which means male's metabolic minutes change might have a bigger impact on the BMI change.


#The relationship between BMI change and shift

```{r}
ggplot(data_complete_cases, aes(x=shift, y=BMI_change)) + 
    geom_boxplot() + xlab("shift") + ylab("change in BMI") + ggtitle("The relationship between BMI change and shift") + scale_x_discrete(limits = c("7am", "8am", "9am", "10am", "11am", "12pm", "1pm", "2pm", "other"))
```
According to the boxplot of the relationship between BMI change and shift, BMI change in 7am, 8am and 2pm are sighter higher than in other times and it has the lowest distribution of median from 11am to 1pm. This suggests that shift is an influence factor of BMI change.



#The relationship between BMI change and metabolic

```{r}
ggplot(data_complete_cases, aes(x=log(Total_met_min_imp), y=log(BMI_change))) +geom_point()+ xlab("log(Total metabolic minutes)") + geom_smooth(method = "lm", se = FALSE)+ ylab("log(Change in BMI)") + ggtitle("The relationship between BMI Change and Total Metabolic Minutes")
```
According to the scatterplot of the relationship between BMI change and Total metabolic minutes after log both of the variables, it shows a tendency of homoskedasticity starting from Total metabolic minutes 5. There are fewer points distributed in Total metabolic minutes less than 5 so it does not have much effects on the relationship. Overall it shows a constant variance.



#The relationship between BMI change and gender
```{r}
ggplot(data_complete_cases, aes(x=gender, y=BMI_change)) + 
    geom_boxplot(outlier.shape = NA) + xlab("gender") + ylab("change in BMI") + ggtitle("The relationship between BMI change and gender") 
```
According to the outpit of the boxplot, it suggests that BMI change are not significantly different for male and female. Female has a similar median value as male, so gender might not have much effect on the change in BMI.




### Statistical Modelling

### Baseline model
To examine the effects of shift and total metabolic minutes on the empolyees' weight gain, we first include a multiple linear regression model into our analysis. Based on our findings in the exploratory data analysis, we decided to use logrithm transformation for both the response variable "BMI change" and the explanatory variable "Total metabolic minutes" to fit the linear relationship. If the variable "Total metabolic minutes" has a value of zero, a small constant 0.01 will be added to the value in order to perform the log transformation. Other predictors included in the baseline model are: shift, age, and gender. The specific model we use is as follows: 


```{r}
data_complete_cases$ttm = data_complete_cases$Total_met_min_imp + 0.01
m1 = lm(log(BMI_change) ~ log(ttm) + shift + Age + gender, data = data_complete_cases)
summary(m1)

# RSS
sum(residuals(m1)^2)
```



Based on a significance level of 0.05 ($\alpha = 0.05$), the estimation results show that the total metabolic minutes has no statistically significant effect on the increase of BMI. The shift time also do not have significant effects on the change in BMI, except for the shift at 8 am. Specifically, for the employees whose shift were at 8 am, their average increase in the BMI is 1.36 higher than the empoyees at other shift time ($\exp(0.31) \approx 1.36$). The model has a residual sum of squares (RSS) of 67.6, with a R-squared of 0.13, suggesting that about 13\% of the variation in the BMI change can be explained by the model.  



### T-test 
T-test to see if there is any significant difference between male and female
```{r}
t.test(BMI_change ~ gender, data = data_complete_cases )
t.test(Total_met_min_imp ~ gender, data = data_complete_cases )
```

```{r}
# export regerssion output to latex
# install.packages("texreg")
# library(texreg)
# texreg(list(m1,m2))
```


### model with interaction terms
In the exploratory data analysis, we observe that the effect of total metabolic minutes on BMI change appears to be different for male and female workers (Figure xxx). In this subsection, we include the interaction term between total metabolic minutes and gender to our baseline model to examine whether the difference is statistically significant. 

```{r}
data_complete_cases$ttm = data_complete_cases$Total_met_min_imp + 0.1
m2 = lm(log(BMI_change) ~ log(ttm)*gender + shift + Age, data = data_complete_cases)
summary(m2)
```



```{r}
# F-test
anova(m1, m2)
```

The results suggest that the effects of total metabolic minutes on BMI change are not significantly different for male and female workers. The results from the F-test also confirms that including the intersection term between total matebolic minutes and gender does not significantly improve the model (p-value = 0.12). 



### Model diagnosis
### Use qqplot to check for normality
```{r}
plot(m1, 2)
```

The qqplot is very close to a straight line, indicating the normality assumption has been satisfied. 



## use the residual plot to check for homosckedasticity
```{r}
# m1_residual <- resid(m1)
# m1_fitted <- fitted(m1)
# plot(m1_fitted,m1_residual, xlab = "fitted values", ylab = "residuals", main = "residual plot for model 1")


plot(m1,1)
```

The residual plot shows a slight tendency of increasing vriance with the fitted values, suggesting the homesckedascity assumption of our linear model might be violated. To address this issue, we use the robust sandwich variance estimator for our model 1. Table xxx compares the original standard errors with the sandwich standard errors. 


```{r}
library(alr4)
#hccm(m1, type="hc3")

cbind("Original SE"=sqrt(diag(vcov(m1))), "Sandwich SE"=sqrt(diag(hccm(m1, type="hc3"))))
```

The sandwich estimation does not change the standard errors much, and our main findings still holds. That is, the total metabolic minutes do not have significant effects on the increase in BMI over the 8 month study period. In addition, only the shift at 8 am is associated with the increase in BMI. 





